---
output:
  html_document:
    css: styles.css
    self_contained: true
---

# Data Analysis + Visualization

<br>



#### I am currently working towards a minor in Data in Society at UC Davis, which has already taught me a great deal of applicable skills in communications. 
<br>

#### In addition to being able to create complex and thorough data visualizations programmed in R, **I can also tell persuasive stories through the data itself.**
<br>

#### The way I see it, the world of communications has so much more to do with data than most would assume. From understanding demographic information of key publics to using data in marketing materials, I believe my skills in data interpretation and representation provide unique value to communications work.
<br>


#### Below are a few examples of data visualizations I have created recently...


<br>





### Interactive Barcharts:
#### *(Move your mouse over them!)*

<br>

```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
library(ipumsr)
library(maps)
library(broom)
library(mapproj)
library(leaflet)
library(sf)
library(tigris)
library(htmltools)
library(htmlwidgets)
#labproj_extract <- define_extract_micro(collection = "usa", description = "Group 8 Lab Data (For Bar)",
     #samples = c("us1990b"),
    #variables = c("YEAR", "STATEFIP", "PERWT",
                  #"BPL", "BPLD", "INCWAGE", "EDUCD"))
#labproj_submitted <- submit_extract(labproj_extract)
#labproj_complete <- wait_for_extract(labproj_submitted)
#"labproj" <- download_extract(labproj_submitted)

#labproj <- read_ipums_ddi("usa_00021.xml") %>% read_ipums_micro()
#Save as RDS
#saveRDS(labproj, "labproj-bar.RDS")

lab_projbar <- readRDS("data/labproj-bar.RDS")

#Convert the labeled columns into numbers
lab_projbar <- lab_projbar %>%
  mutate(across(where(~inherits(., "haven_labelled")), ~as.numeric(.)))


lab_projbar <- lab_projbar %>% 
  filter(YEAR == 1990 & !BPL %in% c(1:99) & INCWAGE > 0 & INCWAGE < 999998 & !is.na(INCWAGE)) %>% 
  mutate(Birth_place = factor(case_when(BPL %in% c(150:199) ~ "North America",
                                        BPL %in% c(200:299) ~ "North America",
                                        BPL %in% c(300:399) ~ "South America",
                                        BPL %in% c(400:499) ~ "Europe",
                                        BPL %in% c(500:599) ~ "Asia",
                                        BPL == 600 ~ "Africa",
                                        BPL %in% c(700:710) ~ "Oceania",
                                        BPL == 800 ~ "Other/Unknown",
                                        TRUE ~ "Other/Unknown")))


format_currency <- function(x) {
  ifelse(is.na(x), "No data", paste0("$", format(round(x, 0), big.mark = ",")))
}



lab_projbar <- lab_projbar %>% 
  filter(YEAR == 1990 & !BPL %in% c(1:99) & EDUCD > 1 & INCWAGE < 999998 & !is.na(INCWAGE)) %>% 
  mutate(EDULVL = factor(case_when(EDUCD %in% c(2:61) ~ "No HS Diploma",
                                        EDUCD %in% c(62:64) ~ "High School Diploma",
                                        EDUCD == 71 ~ "Some College (No Degree)",
                                        EDUCD == 101 ~ "Bachelor's degree",
                                        EDUCD == 114 ~ "Master's degree",
                                        EDUCD == 116 ~ "Doctoral degree",
                                        TRUE ~ NA)))

library(ghibli)
pal <- ghibli_palette(name = "MarnieMedium2", n = 7, type = "discrete")


pal2 <- c("#09122C", "#4C3A51", "#BE3144", "#8E1616", "#E17564", "#E7AB79", "#B46060")


#install.packages("showtext")
library(showtext)
font_add_google("Baskervville SC", "baker", db_cache = FALSE)
font_add_google("IBM Plex Sans", "ibm", db_cache = FALSE)
font_add_google("Special Elite", "elite", db_cache = FALSE)
showtext_auto()

#install.packages("ggiraph")
library(ggiraph)




bar1 = lab_projbar %>%
  drop_na(EDULVL, Birth_place) %>%
  group_by(EDULVL, Birth_place) %>%
  summarise(avg_income = weighted.mean(INCWAGE, w = PERWT, na.rm = TRUE), .groups = "drop") %>%
  group_by(EDULVL) %>%
  mutate(total_income = sum(avg_income)) %>%
  ungroup() %>%
  ggplot(aes(x = reorder(EDULVL, total_income), y = avg_income, fill = Birth_place,
             tooltip = paste0(Birth_place, "\nAvg Income: $", format(round(avg_income, 0), big.mark = ",")), data_id = Birth_place)) +
  geom_col_interactive() +
  scale_y_continuous(labels = scales::dollar_format()) +
  scale_fill_manual(values = pal) +
  labs(title = "Income by Education Level, Seperated by Birthplace (1990)", subtitle = "Data Source: U.S. Census Bureau (1990)", x = "Education Level", y = "Average Income", , fill = "Birth Place") +
  theme_minimal() +
  theme(text = element_text(family = "ibm", size = 16)) +
  theme(plot.title = element_text(family = "baker")) +
  theme(plot.subtitle = element_text(family = "baker")) +
  theme(plot.title = element_text(size = 24))




portfolio1 = girafe(ggobj = bar1, width_svg = 17, height_svg = 10.2)




```


```{r echo=FALSE}

portfolio1

```




```{r echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}




library("readxl")

tops = read_xlsx("data/Billboard Hot 100 Number Ones Database.xlsx", sheet = 2)

#Adding a year column 
tops$Year = year(tops$Date)

tops$Decade = factor(case_when(tops$Year %in% c(1960:1969) ~ "60s",
                     tops$Year %in% c(1970:1979)~ "70s",
                     tops$Year %in% c(1980:1989)~ "80s",
                     tops$Year %in% c(1990:1999)~ "90s",
                     tops$Year %in% c(2000:2009)~ "2000s",
                     tops$Year %in% c(2010:2019)~ "2010s",
                     tops$Year %in% c(2020:2024)~ "2020s", TRUE ~ NA),
  levels = c("60s", "70s", "80s", "90s", "2000s", "2010s", "2020s")
)

#tops$Decade2 = floor(tops$Year / 10) * 10

yearly_rap_verse = aggregate(tops$`Rap Verse in a Non-Rap Song` ~ Year, data = tops, 
          FUN = function(x) sum(x == 1))

tops$Explicit = as.numeric(tops$Explicit)
tops$Genre = str_split(tops$`CDR Genre`, ";") %>% map_chr(1)


#Same for explicit lyrics

explicit_over_time = aggregate(tops$Explicit ~ Year, data = tops,
                               FUN = function(x) sum(x ==1))


#Visualizing it with a density plot, added scales for x and y to clearly show which decade is which
explicit_bar1 = ggplot(explicit_over_time, aes(x = Year, y = `tops$Explicit`, 
                               tooltip = paste0(Year, ":", format(round(`tops$Explicit`, 0), big.mark = ",")), data_id = Year)) +
        geom_bar_interactive(stat = "identity", fill = "darkred") +
        scale_y_continuous(breaks = seq(0, 15, by = 1)) +
        scale_x_continuous(breaks = seq(1950, 2025, by = 10)) +
        labs(title = "Number of Explicit Billboard #1 Songs per Year (1960-2024)", subtitle = "Data Source: Chris Dalla Riva",
        x = "Year", y = NULL) +
        theme_minimal() +
        theme(text = element_text(family = "ibm", size = 12)) +
        theme(plot.title = element_text(family = "baker")) +
        theme(plot.subtitle = element_text(family = "baker")) +
        theme(plot.title = element_text(size = 17))


explicit_bar = girafe(
  ggobj = explicit_bar1,
  width_svg = 9, height_svg = 5,
  options = list(
    opts_hover(css = "fill:darkred;"),
    opts_hover_inv(css = "opacity:0.7;")
  )
)


  
  

rexplbygenre = tops %>%
  filter(Explicit == 1, !is.na(Genre), !is.na(Year)) %>%
  group_by(Year, Genre) %>%
  summarize(Count = n(), 
         Songs = paste(Song, collapse = "\n"),
         .groups = "drop") %>%
  ggplot(aes(x = Year, y = Count, color = Genre, fill = Genre,
     tooltip = paste0(Genre, " (", Year, "): ", Count, " explicit #1s:\n \n", Songs), data_id = paste0(Genre, Year))) +
  geom_col_interactive() +
  scale_y_continuous(breaks = seq(0, 15, by = 1)) +
  scale_x_continuous(breaks = seq(1950, 2020, by = 10)) +
  scale_fill_manual(values = pal2) +
  scale_color_manual(values = pal2) +
  theme_minimal() +
  labs(title = "Billboard #1 Hit Songs with Explicit Lyrics, by Genre (1960-2019)", subtitle = "Data Source: Chris Dalla Riva",
        x = "Year", y = NULL, fill = "Genre", color = "Genre") +
        theme_minimal() +
        theme(text = element_text(family = "ibm", size = 12)) +
        theme(plot.title = element_text(family = "elite")) +
        theme(plot.subtitle = element_text(family = "ibm")) +
        theme(plot.title = element_text(size = 17))







portfolio2 = girafe(
  ggobj = rexplbygenre,
  width_svg = 9, height_svg = 5,
  options = list(
    opts_hover(css = "opacity:1; fill-opacity:1;stroke:none"),
    opts_hover_inv(css = "opacity:0.7; fill-opacity:0.7;stroke:none"),
    opts_tooltip(css = "background:black;color:white;padding:6px;font-family:'Special Elite';")
  )
)

```

```{r echo=FALSE}

portfolio2

```